---
title: Starting R 
author: R Pruim & D Kaplan
date: "Data Computing Instructors' Workshop"
output: 
  ioslides_presentation:
    widescreen: true
---


<style type="text/css">
span.boxed {
  border:5px solid gray;
  border-radius:10px;
  padding: 5px;
}
span.invboxed {
  border:5px solid gray;
  padding: 5px;
  border-radius:10px;
  color: white;
}
table, td, th
{
border:0px;
}
</style>


```{r include=FALSE}
require(mosaic)
require(ggplot2)
```


## Modeling 

Modeling is really the starting point for the `mosaic` 
design.

 * linear models (`lm()` and `glm()`) defined the template
 * lattice graphics use the template (so we chose lattice)
 * we added functionality so numerical summaries 
 can be done with the same template
 * additional things added to make modeling easier for beginners.

## Models as Functions

```{r}
model <- lm(width ~ length * sex, 
            data=KidsFeet)
Width <- makeFun(model)
Width( length=25, sex="B")
Width( length=25, sex="G")
```

## Models as Functions -- Plotting

```{r, include=FALSE}
trellis.par.set(
  superpose.symbol=list(col=c('navy','red'), pch=16), 
  superpose.line=list(lty=1, col=c('navy','red'))
)
```
                                                       
```{r, fig.keep='last'}
xyplot( width ~ length, data=KidsFeet, 
        groups=sex, auto.key=TRUE )
plotFun( Width(length, sex="B") ~ length, 
         col=1, add=TRUE)
plotFun( Width(length, sex="G") ~ length, 
         col=2, add=TRUE)
```

```{r, include=FALSE}
trellis.par.set(theme=col.mosaic())
```

## Resampling -- You can do() it!


## Lady Tasting Tea

 * Often used on first day of class
 
 * Story
 
   * woman claims she can tell whether milk
   has been poured into tea or vice versa.
   
   * Question: How do we test this claim?
   
```{r echo=FALSE }
require(mosaic)
trellis.par.set(theme=col.mosaic())
require(knitr)
opts_chunk$set(size='small', cache=TRUE)
options(width=90)
set.seed(12345)
```

```{r echo=FALSE}
require(knitr)
opts_chunk$set( 
  fig.width=8, 
  fig.height=3,
  out.width=800,
  out.height=300,
  fig.align="center"
  )
```

## Computer Simulation

Use `rflip()` to simulate flipping coins
```{r}
rflip()
```

## Computer Simulation

Faster if we flip multiple coins at once:
```{r} 
rflip(10)
```
 * easier to consider `heads` = correct; `tails` = incorrect than to compare with a given pattern
   * this switch bothers me more than it bothers my students

## Let's do that a lot of times

`rflip(10)` simulates 1 lady tasting 10 cups 1 time.

We can do that many times to see how guessing ladies do:
```{r}
do(2) * rflip(10)
```

 * `do()` is clever about what it remembers
 * 2 isn't many -- we'll do many next.

## The distribution of guessing ladies


```{r fig.width=8, fig.height=3, out.width=800, out.height=300}
Ladies <- do(5000) * rflip(10)
head(Ladies, 1)
histogram( ~ heads, data=Ladies, width=1 )
```

## How often does guessing score 9 or 10?


```{r}
tally( ~(heads >= 9) , data=Ladies)
```

## How often does guessing score 9 or 10?


```{r}
tally( ~(heads >= 9) , data=Ladies, format="prop")
```

## How often does guessing score 9 or 10?


```{r}
tally( ~(heads >= 9) , data=Ladies, format="prop")
 prop( ~(heads >= 9), data=Ladies)
```

## A general approach to randomization


 1. Do it for your data
 2. Do it for "random" data
 3. Do it lots of times for "random" data
 
 * definition of "random" is important, but can
 often be handled by `shuffle()` or `resample()`

## Example: Do mean ages differ by sex?

```{r}
diffmean(age ~ sex, data=HELPrct)
do(1) * 
  diffmean(age ~ shuffle(sex), data=HELPrct)
Null <- do(5000) * 
  diffmean(age ~ shuffle(sex), data=HELPrct)
```

## Example: Do mean ages differ by sex?


```{r fig.width=8, fig.height=3, out.width=800, out.height=300}
prop( ~(abs(diffmean) > 0.7841), data=Null )
histogram(~ diffmean, data=Null, v=-.7841) 
```

## Example:  Bootstrap for difference in means


```{r fig.width=8, fig.height=3, out.width=800, out.height=300}
Bootstrap <- do(5000) * 
  diffmean(age~sex, data= resample(HELPrct))

histogram( ~diffmean, data=Bootstrap, 
                      v=-.7841, glwd=4 )
```

## Example:  CI for difference in mean ages

```{r}
cdata(~diffmean, data=Bootstrap, p=.95)
confint(Bootstrap, method="quantile")
```

## Example:  CI for difference in mean ages

```{r}
confint(Bootstrap)  # default uses st. err
```

## Randomization and linear models


```{r}
do(1) * lm(width ~ length, data=KidsFeet)
do(3) * lm( width ~ shuffle(length), data=KidsFeet)
```

## Randomization and linear models


```{r}
do(1) * 
  lm(width ~ length + sex, data=KidsFeet)
do(3) * 
  lm( width ~ length + shuffle(sex), 
                       data=KidsFeet)
```


## Randomization and linear models

```{r fig.width=8, fig.height=3, out.width=800, out.height=300, cache=TRUE}
Null <- do(5000) * 
  lm( width ~ length + shuffle(sex), 
                       data=KidsFeet)
histogram( ~ sex.G, data=Null, 
           v=-0.2325, glwd=4)
```

## Randomization and linear models

```{r fig.width=8, fig.height=3, out.width=800, out.height=300}
histogram(~ sex.G, data=Null, 
           v=-0.2325, glwd=4)
prop(~ (sex.G <= -0.2325), data=Null)
```


