# Total vs Partial Change

```{r include=FALSE}
require(mosaic) # leave this chunk alone
knitr::opts_chunk$set(dev="svg",out.width="50%")
```


## Car Prices

The file `http://tiny.cc/mosaic/used-hondas.csv` contains data collected by students in Macalester's "Introduction to Statistical Modeling" course about asking prices for used Hondas.

```{r cache=TRUE}
cars = read.csv("http://tiny.cc/mosaic/used-hondas.csv")
names(cars) # the variable names
```

### Several models to fit

For each of these, fit the model with the `lm()` function: `lm(formula, data = cars)`. Call the models you create `mod1`, `mod2`, `mod3`, and `mod4`.

1. `Price ~ Age`
2. `Price ~ Mileage`
3. `Price ~ Age + Mileage`
4. `Price ~ Age * Mileage`

```{r echo=FALSE}
mod3 <- lm(Price ~ Age + Mileage, data = cars)
```

### Calculating model values

Extract the model function from each model with a command like this:
```{r}
f3 <- makeFun(mod3)
```

For each of the models, find the model price for a car that's 7 years old with 80,000 miles.

## Differences and derivatives

You can quantify the impact of each variable by comparing two cars with different values for that variable.  For example:
```{r}
f3(Age = 8, Mileage = 80000) - f3(Age = 7, Mileage = 80000)
```

Or, if you prefer, you can compute the derivative function:
```{r warning=FALSE}
df3_dAge <- D(f3(Age = Age, Mileage = Mileage) ~ Age)
df3_dAge(Age = 7, Mileage=80000)
```

## Tasks

For each of the four models, 

* Calculate the **total change** for a car that ages from 4 to 5 years while the mileage increases from 40,000 to 50,000.
* Calculate the **partial change** with respect to `Age` for a car that ages from 4 to 5 years with the mileage staying at 40,000. 
* Calculate the **partial change** with respect to `Mileage` for a car aged 4 years with 40,000 miles.

Explain why the four models give different results.

## Confidence intervals

When evaluating the model functions, you can get confidence or prediction intervals by adding an argument `interval = "confidence"` or, alternatively, `interval = "prediction"`.

## Regression tables

The standard regression tables can be computed using the `summary()` function on the model. (That is, on `mod3`, not on `f3`.) Compare the regression table results to your partial and total differences.

## Going further

Economists often use the logarithm of price rather than the price itself.  You can do this simply, e.g.:

```{r}
mod3a <- lm( log10(Price) ~ Age + Mileage, data=cars)
f3a <- makeFun(mod3a)
```

How would you determine whether using a log transform like this would produce a "better" model?


### Interpreting the coefficients

For models 1 through 3, it's comparatively easy 

### Price versus age and mileage

Here are two models of price.  For each, the function from the model has been explicitly 

## Housing Prices

Using the `SaratogaHouses.csv` data, estimate the partial change of adding a fireplace onto a house of 2000 square feet living area, 3 bedrooms, and 2 baths.

```{r houses, cache=TRUE}
houses <- read.csv("http://tiny.cc/mosaic/SaratogaHouses.csv")
names(houses)
```

Make models both with and without interactions between living area and fireplace. (You can also put in bedrooms and baths, if you like.) 

* How much does the partial change of a fireplace depend on the particular model? 



## Wages

What's the value of a college education? (Or, what was it in the decades before 1985, when the `CPS85` data were collected?)

Find both the **partial** and the **total change** of extending education from 12 years to 16 years in this model

```{r}
wmod1 <- lm(wage ~ educ*sector + educ*exper, data=CPS85)
w1 <- makeFun(wmod1)
```

* You'll have to choose a sector and an amount of experience to create a number for the partial change.  Do so.

* As `educ` changes, so does `exper`, since the year of education comes at the cost of a reduced amount of experience.  You might also reasonably infer a change in sector due to education.
    * What's the total change of `educ` when balanced against a corresponding change in `exper`, while holding `sector` constant?
    * How much does a plausible change in `sector` change things?





